---
title: "Geospatial Analysis"
author: "Ahmad Ilu"
date: "August, 2024"
output: beamer_presentation
mainfont: "Open Sans" # Try out some font options if xelatex
theme: Madrid
subtitle: Nigeria's Geospatial at a Glance
institute:
 - Iluahmad9@gmail.com
 - Economist/ Policy Researcher
---
## Key Takeaways
The Data used in this brief Geospatial note are sourced from from [Wikipedia](https://en.wikipedia.org/wiki/States_of_Nigeria), [NBS(2019)](file:///C:/Users/Ahmad%20Ilu/Downloads/2019%20WASH%20NORM%20SURVEY%20FINDINGS%20-%20INFOGRAPHIC%20SUMMARY.pdf) and [Federal Ministry of Health](https://hfr.health.gov.ng/statistics/tables). Also the codes and data can be accessed on my [Github](https://github.com/Iluahmad/Academic-Profile) page.

-   only 9% Have access to basic water,sanitation, hygiene services
-   70% Use basic drinking water services ( People living in Lagos are three times more likely to use basic water supply services than those living in Taraba)
- 9 liters Average per capita volume of water per day for the rural population
-   46 million  Practice open defecation
- Cases of Open defecation is more prevalent in Kwara State at $\%64$ compared to $\%2$ in Abia and Zamfara States and also higher than National Average of $\%23$.
- Almost half of Jigawa population has access to basic hygiene service compared to only 1% of citizens in Ebonyi state.

## Flash!
This is a rough sketch of Nigeria's map structure.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(sf)
library(ggplot2)
library(rnaturalearth)
library(rnaturalearthdata)
library(dplyr)
library(tmap)
library(readxl)
```

```{r echo=FALSE}
Nigerian_States <- read_excel("C:/Users/Ahmad Ilu/Downloads/Nigerian States.xlsx", sheet = "Sheet1")
```

```{r echo=FALSE}
View(Nigerian_States)
```

```{r include=FALSE}
#Loading the Shape Data
  ng_sf <- st_read(dsn = "C:/Users/Ahmad Ilu/Downloads/NGA_State_Boundaries_V2_-8623244326519598647/grid3_nga_boundary_vaccstates.shp")
```

```{r echo=FALSE}
ng_sf %>% ggplot() + geom_sf()+ ggthemes::theme_map()
```
```{r include=FALSE}
# Wrangling the Data
merged_data = ng_sf %>% left_join(Nigerian_States, by = c( 'statename'))
merged_data 
```

## Nigeria's Population
```{r echo=FALSE}
merged_data %>% ggplot()+ geom_sf(aes(fill = Population/1e6), color = 'white', lwd = 0.3)+ labs(title ="Nigerian Geospatial", fill = "Population\n(millions)")+ scale_colour_viridis_d() + ggthemes::theme_map()+ theme(legend.position = 'top', legend.title = element_text(face = "bold"), plot.title = element_text(size = 50,color = "steelblue", hjust = 0.5, face = "bold" ) )

```

## Slide with Access to Basic Water %
```{r echo=FALSE}
# Access to Basic Water %
merged_data %>% 
  tm_shape()+
  tm_polygons("Access to Water %", 
              palette = c( "#347fc1", "#abdea3","blue","steelblue","#feaf62"), 
              title = "Access to Basic Water %",
              legend.format = list(text.separator = "-"))+ tm_text( "statename", 
                                                                    col = "black",
                                                                    #          auto.placement = TRUE,
                                                                    size = .5)+
  tm_layout(
    main.title = "Access to Basic Water %",
    legend.position = c("right", "bottom"))+
  tm_scale_bar(breaks = c(0,25,50), text.size = .8, position = c("left","bottom"))+
  tm_credits(text = "Source: NBS 2019", size = 0.6, fontface = "bold", position = c("right","top"))+
  tm_compass(type = "4star",size = 1,
             position = c("left","top"), 
             cardinal.directions = c('North'),
             bg.color = 'red',
             bg.alpha = .13)

```

## Access to Basic Hygiene %
```{r echo=FALSE}
# Access to Basic Hygiene %
merged_data %>% 
  tm_shape()+
  tm_polygons("Access to Basic Hygiene %", 
              palette = c( "#666666","#B1DAC5","#7570B3","#71B966","#66A61E" ), 
              title = "Access to Basic Hygiene %",
              legend.format = list(text.separator = "-"))+ tm_text( "statename", 
                                                                    col = "black",
                                                                    #          auto.placement = TRUE,
                                                                    size = .5)+
  tm_layout(
    main.title = "Access to Basic Hygiene %",
    legend.position = c("right", "bottom"))+
  tm_scale_bar(breaks = c(0,25,50), text.size = .8, position = c("left","bottom"))+
  tm_credits(text = "Source: NBS 2019", size = 0.6, fontface = "bold", position = c("right","top"))+
  tm_compass(type = "4star",size = 1,
             position = c("left","top"), 
             cardinal.directions = c('North'),
             bg.color = 'red',
             bg.alpha = .13)
```


## Health Care Facility in Nigeria

```{r}
merged_data %>% 
  tm_shape()+
  tm_polygons("Health Care Facility", 
              palette = c("skyblue" ,"lightblue","steelblue","#2E5B88","#1B9E77" ), 
              title = "No of health care Facilities",
              legend.format = list(text.separator = "-"))+ tm_text( "statename", 
                                                                    col = "black",
                                                                    #          auto.placement = TRUE,
                                                                    size = .5)+
  tm_layout(
    main.title = "Health care Facilities across States",
    legend.position = c("right", "bottom"))+
  tm_scale_bar(breaks = c(0,25,50), text.size = .8, position = c("left","bottom"))+
  tm_credits(text = "Source: FMH 2019", size = 0.6, fontface = "bold", position = c("right","top"))+
  tm_compass(type = "4star",size = 1,
             position = c("left","top"), 
             cardinal.directions = c('North'),
             bg.color = 'red',
             bg.alpha = .13)
```

